The Origins of Congestion\\and Network-assisted end-to-end\\Congestion Control

### 2020-07-09 slides/slides.tex
Final Version

	\only<1-2>{\begin{column}{.45\textwidth}\end{column}}
	\only<3->{
	\begin{column}{.45\textwidth}
		\begin{itemize}
		\item Links have limitations.
		\item<4-> Data is split up into smaller packets.
		\item<5-> Network nodes contain buffers storing packets.
		\end{itemize}
		\only<5->{
		\begin{alertblock}{}
		Origins of Congestion are versatile.
		\end{alertblock}
		}
	\end{column}
	}
\end{columns}
\end{frame}

@@

\subsection{Buffers}
\label{sec:buffers}
\begin{frame}{Buffers}
~\\
\begin{block}{Purpose}
Buffers store packets in network nodes waiting for transmission.
\end{block}
~\\
\only<2->{
Sizing Buffers:
\begin{itemize}
\item Buffer size should be large enough to handle packet bursts.
\item Too large buffers result in an increase of queuing delay and \textit{over-buffering}.
\end{itemize}
}
~\\
\only<3->{
\begin{enumerate}
\item[$\Rightarrow$] Buffers should be designed according to their application environment.
\end{enumerate}
}
\end{frame}

\begin{frame}{Bufferbloat}
\begin{itemize}
\item Refers to frequently full, excessively large buffers.
\item<2-> Queued packets will not be dropped, resulting in:
\only<2->{
	\begin{itemize}
	\item Unnecessary network delay (\textit{latency}) and packet delay fluctuation (\textit{jitter}).
	\item Increased congestion, as TCP is not being notified of congestion.
	\item Decreases network throughput.
	\end{itemize}
}
\item<3-> With memory getting cheaper over the years, buffers saw an increase in size.
\end{itemize}
\end{frame}

@@

		\draw[-, dotted] (net1) -- (Router1);
		\draw[-] (Router1) -- (Router2);
		\draw[-] (Router2) -- (Router3);
		\draw[-, dotted] (Router3) -- (net2);
	\end{tikzpicture}
\end{figure}
~\\
\begin{minipage}{\textwidth}
\begin{columns}
%Content
\only<2->{
	\begin{column}{0.5\textwidth}
		\begin{block}{Transmission Delay}
			Depends on bandwidth and length of the connection.
		\end{block}
	\end{column}
}
\only<3->{
	\begin{column}{0.5\textwidth}
		\begin{block}{Processing Delay}
			Time it takes to process a single packet.
		\end{block}
	\end{column}
}
%Placeholders
\only<1>{\begin{column}{0.5\textwidth}\end{column}}
\only<1-2>{\begin{column}{0.5\textwidth}\end{column}}
\end{columns}
\end{minipage}

\begin{minipage}{\textwidth}
\begin{columns}
\only<4->{
	\begin{column}{0.5\textwidth}
		\begin{block}{Inter-Packet Delay}
			Allows the receiver to prepare for transmission.
		\end{block}
	\end{column}
}
\only<5->{
	\begin{column}{0.5\textwidth}
		\begin{block}{Queuing Delay}
			Time a packet waits in the buffer waiting for transmission.
		\end{block}
	\end{column}
}
\only<1-3>{\begin{column}{0.5\textwidth}\end{column}}
\only<1-4>{\begin{column}{0.5\textwidth}\end{column}}
\end{columns}
\end{minipage}
\end{frame}

@@

\begin{column}{0.5\textwidth}
	\begin{block}{Delay}
		\begin{itemize}
		\only<2->{\pgfsetfillopacity{0.2}}
		\item Transmission Delay
		\item Processing Delay
		\item Inter-Packet Delay
		\only<2->{\pgfsetfillopacity{1}}
		\item Queuing Delay
		\end{itemize}
	\end{block}
\end{column}
\end{columns}
~\\~\\
\only<2->{
\begin{itemize}
\item Congestion control algorithms can improve queuing delay, number of queued packets and buffer management.
\end{itemize}
}
\end{frame}

@@

\begin{frame}{TCP}
~\\
\begin{block}{Transmission Control Protocol}
\begin{itemize}
\item Lives on transport layer, communicates with application layer via sockets ($\rightarrow$ ports) and uses IP as internet protocol ($\rightarrow$ TCP/IP).
\item Ensures successful transmission of all data.
\end{itemize}
\end{block}
\begin{itemize}
\item<2-> Transmitted data needs to be acknowledged.
\item<2-> Adjusts to the current state of congestion.
\item<3-> Many different implementations exist.
\end{itemize}
\end{frame}

\begin{frame}[fragile]{TCP Header}
\centering
\begin{figure}
\centering
\begin{bytefield}[bitheight=2.2\baselineskip, bitwidth=0.635\baselineskip]{32}
\bitbox{16}{Source Port} & \bitbox{16}{Destination Port} \\
\bitbox{32}{Sequence Number} \\
\bitbox{32}{Acknowledgment Number} \\
\bitbox{4}{Data Offset} & \bitbox{4}{Rsrvd} & \bitbox{1}{\tiny C\\W\\R} & \bitbox{1}{\tiny E\\C\\E} & \bitbox{1}{\tiny U\\R\\G} & \bitbox{1}{\tiny A\\C\\K} & \bitbox{1}{\tiny P\\S\\H} & \bitbox{1}{\tiny R\\S\\T} & \bitbox{1}{\tiny S\\Y\\N} & \bitbox{1}{\tiny F\\I\\N} & \bitbox{16}{Window Size} \\
\bitbox{16}{Checksum} & \bitbox{16}{Urgent Pointer} \\
\bitbox{32}{Options} \\
\bitbox{32}{Data}
\end{bytefield}
\end{figure}
\end{frame}

@@

\begin{frame}{Perceiving Congestion}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{block}{Round-trip Time (\textit{RTT})}
Describes the amount of time between sending a data packet and receiving its acknowledgment packet.
\end{block}
\end{column}
\begin{column}{0.5\textwidth}
\begin{block}{Retransmission Timeout (\textit{RTO})}
The maximum amount of time the sender will wait for the acknowledgment of a packet.
\end{block}
\end{column}
\end{columns}
~\\
\only<2->{
\begin{itemize}
\item Classic TCP perceives congestion by detecting packet loss.
\item Other implementations also perceive congestion by determining variations in RTT over time.
\end{itemize}
}
\end{frame}

@@

	\only<1->{\path (s1) edge [above] node {\tiny A} (r1);}
	\only<2->{\path (s2) edge [dashed, above] node {\tiny B} (loss);}
	\only<3->{\path (s3) edge [below] node {\tiny C} (r3);}
	\only<4->{\path (r4) edge [above] node {\tiny A} (s4);}
	\only<5->{\path (r6) edge [above] node {\tiny A} (s6);}

@@

\subsection*{Adjusting for Congestion}
\label{sec:ac}
\begin{frame}{Adjusting for Congestion}
~\\
\begin{block}{Goal}
\begin{itemize}
\item Do not fill up buffers in the network.
\item Achieve and hold maximum throughput.
\item Adapt to the current state of congestion as quickly as possible.
\end{itemize}
\end{block}
\only<2->{
\begin{itemize}
\item CA algorithms are based on Additive-Increase Multiplicative-Decrease principle.
\item Variable $cwnd$ (congestion window) describes the maximum amount of data sent before ACKs are needed.
\end{itemize}
}
\end{frame}

\begin{frame}{Adaption - Additive Increase}
~\\
\begin{block}{Formula}
$cwnd \coloneqq cwnd + u$
\end{block}
\begin{itemize}
\item Increases the data rate slowly to reach the networks maximum throughput.
\item Prevents the rapid filling of buffers.
\item Proposed value for $u$: $\frac{1}{cwnd}$
\end{itemize}
\end{frame}

\begin{frame}{Adaption - Multiplicative Decrease}
~\\
\begin{block}{Formula}
$cwnd \coloneqq d \cdot cwnd, 0<d<1$
\end{block}
\begin{itemize}
\item Decreases the data rate drastically.
\item Allows the nodes to recover and buffer lengths to decrease.
\item Proposed value for $d$: $\frac{1}{2}$
\end{itemize}
\end{frame}

\begin{frame}{Adaption - Slow-Start}
~\\
\begin{block}{Formula}
$cwnd \coloneqq cwnd + 1$ for each received ACK packet.
\\$\Rightarrow$ Exponential increase over time.
\end{block}
\begin{itemize}
\item Drastically increases the protocols performance.
\item Used after connection is established or after multiple packet losses have been detected.
\item Switches off after a threshold has been exceeded or congestion has been perceived.
\end{itemize}
\end{frame}

@@

\begin{frame}{Network-assisted Congestion Control}
\begin{itemize}
\item Congestion Detection algorithms are very implicit:
	\begin{itemize}
	\item Packet loss may also have other causes
	\item Methods are inaccurate, and do not inform about the degree of congestion.
	\end{itemize}
\item[$\Rightarrow$] Creates unnecessary network delay.
~\\~\\
\item<2-> Explicit congestion notification is crucial to keep the network stable.
\item[$\Rightarrow$]<2-> Modify the IP header to include a \textit{Congestion Experienced} Flag (CE).
\item<3-> \textit{Explicit congestion notification} (ECN), an extention to TCP/IP, implements this.
\end{itemize}
\end{frame}

\begin{frame}{Principle of Network-assisted Congestion Control}
\begin{center}
\begin{figure}
	\centering
		\begin{tikzpicture}[->, node distance = 2.1cm]
			%network
			\node[client, scale=0.7] (H1) {};
			\node[router, scale=0.7] (R1) [right=of H1] {};
			\node[router, scale=0.7] (R2) [right=of R1] {};
			\node[router, scale=0.7] (R3) [right=of R2] {};
			\node[client, scale=0.7] (H2) [right=of R3] {};
			\path [line] (H1) -- (R1);
			\path [line] (R1) -- (R2);
			\path [line] (R2) -- (R3);
			\path [line] (R3) -- (H2);
			\path [line, dotted] (H2) -- ++(0,-35pt) -| (H1);
			\node[below=0.9cm of R2] (text) [transition] {Congested Signal};

			%packets
			\node[messageclosed,scale=0.5, fill=black!10] (P1) [above right=0.2cm and 1cm of H1] {};
			\node[messageclosed,scale=0.5, fill=black!50] (P2) [above left=0.7cm and 0.3cm of R2] {};
			\node[messageclosed,scale=0.5, fill=black!10] (P3) [above left=0.4cm and 0.2cm of R2] {};
			\node[messageclosed,scale=0.5, fill=black!10] (P4) [above left=0.1cm and 0.1cm of R2] {};
			\node[messageclosed,scale=0.5, fill=black!50] (P5) [above right=0.2cm and 1cm of R3] {};
			\node[above left = 1.5cm and 0cm of R1] (P) [transition] {Packet};
			\node[above right = 1.5cm and -0.3cm of R2] (CN) [transition] {Congested Node};
			\node[above left = 1.5cm and -0.4cm of H2] (MP) [transition] {Marked Packet};
			\path [line] (P) -- (P1);
			\path [line] (CN) -- (R2);
			\path [line] (MP) -- (P5);
		\end{tikzpicture}
\end{figure}
\end{center}
\end{frame}

\subsection{XCP}
\label{sec:xcp}
\begin{frame}{XCP}
~\\
\begin{block}{XCP}
\textit{eXplicit Congestion Protocol} generalizes the ECN proposal.
\end{block}
\begin{itemize}
\item Improves upon ECN by introducing new packet control concepts.
\item Additionally gives feedback about the degree of congestion.
\end{itemize}
\only<2->{
\begin{block}{XCP Header}
\begin{itemize}
\item $H\_cwnd$
\item $H\_rtt$
\item $H\_feedback$
\end{itemize}
\end{block}
}
\end{frame}

\begin{frame}{XCP Sender}
\begin{itemize}
\item As a TCP Sender, the XCP sender also keeps track of $cwnd$ and $RTT$ and deals with packet loss (retransmission).
~\\~\\
\item On data packet depature, the XCP sender will:
	\begin{itemize}
	\item $H\_cwnd \coloneqq cwnd$
	\item $H\_rtt \coloneqq RTT$
	\item $H\_feedback \coloneqq \frac{r \cdot RTT - cwnd}{cwnd}$, where $r$ is the desired data rate.
	\end{itemize}
\item[$\Rightarrow$] The desired data rate can be reached in just one RTT.
~\\~\\
\item On ACK packet arrival, the XCP sender will adjust $cwnd$ according to the received feedback:
	\begin{itemize}
	\item $cwnd \coloneqq max(cwnd + H\_feedback, s)$, where $s$ is the packet size.
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{XCP Router}
\begin{columns}
\begin{column}{.5\textwidth}
\begin{block}{Efficiency Controller (EC)}
\begin{itemize}
\item Maximizes link utilization.
\item Minimizes queue lengths and packet drops.
\item If the node is congested, feedback should be negative, otherwise positive.
\end{itemize}
\end{block}
\end{column}

\begin{column}{.5\textwidth}
\only<2->{
\begin{block}{Fairness Controller (FC)}
\begin{itemize}
\item Divides feedback to packets fairly.
\item If feedback > 0, equal division.
\item If feedback < 0, division in proportion to the packets data rates.
\end{itemize}
\end{block}
}
\end{column}
\end{columns}

\only<3->{
\begin{itemize}
\item Computation of feedback reflecting the current degree of congestion at this node.
\item Allocates bandwidth equally to all XCP senders.
\end{itemize}
}
\end{frame}

\subsection{Evaluation of Congestion Control}
\label{sec:eval}
\begin{frame}{Evaluation of Congestion Control}
Pros:
\begin{itemize}
\item XCP and ECN can easily outperform TCP/IP in certain conditions.
\item Many networks could benefit from using congestion control.
\end{itemize}
\only<2->{
~\\Cons:
\begin{itemize}
\item Many routers run on hardware only $\Rightarrow$ unrecognized packets will be dropped.
\item Network-assisted congestion control often relies on active queue management.
\end{itemize}
}
~\\
\begin{itemize}
\item<3->[$\Rightarrow$] Network-assisted congestion control protocols will not appear widespread in WANs.
\item<4->[$\Rightarrow$] SANs could implement congestion control more easily.
\end{itemize}
\end{frame}

\subsection{DCTP}
\label{sec:dctp}
\begin{frame}{DCTCP}
~\\
\begin{block}{Data center TCP}
\begin{itemize}
\item Extension to TCP for use in data centers.
\item Based on ECN.
\end{itemize}
\end{block}
\begin{itemize}
\item<2-> Tackles the TCP incast problem.
\item<3-> Prevents bursty retransmissions by lowering RTO.
\item<3-> Prevents link under-utilization:
\begin{itemize}
\item <4-> Adjusts to congestion according to the fraction of packets that experienced congestion.
\end{itemize}
~\\~\\
\item<5->[$\Rightarrow$] Implemented in many data centers.
\end{itemize}
\end{frame}

@@

\begin{frame}{Conclusion}
\only<1->{
\begin{block}{Origins of Congestion}
\begin{itemize}
\item Buffer size and Buffer management cause delay and packet loss.
\item Queuing delay results in increased network delay.
\end{itemize}
\end{block}
}
\only<2->{
\begin{block}{Congestion Avoidance}
\begin{itemize}
\item Simple way to tackle congestion.
\item Only gets implicit congestion notification.
\item Often causes instead of resolving network stress.
\end{itemize}
\end{block}
}
\only<3->{
\begin{block}{Network-assisted Congestion Control}
\begin{itemize}
\item Can explicitly control congestion.
\item Results in overall better network performance.
\item Implementations in WANs very challenging, successful in data centers.
\end{itemize}
\end{block}
}
\end{frame}

>>>>

	\only<1-2>{\begin{column}{.45\textwidth}\end{column}}
	\only<3->{
	\begin{column}{.45\textwidth}
		\begin{itemize}
		\item Links have limitations
		\item<4-> Data is split up into smaller packets
		\item<5-> Network nodes contain buffers storing packets
		\end{itemize}
		\only<6->{
		\begin{alertblock}{}
		Origins of Congestion are versatile
		\end{alertblock}
		}
	\end{column}
	}
\end{columns}
\end{frame}

@@

\subsection{Buffers}
\label{sec:buffers}
\begin{frame}{Buffers}
~\\
\begin{block}{Purpose}
Buffers store packets in network nodes waiting for transmission
\end{block}
~\\
\only<2->{
Sizing Buffers:
\begin{itemize}
\item Buffer size should be large enough to handle packet bursts
\item Too large buffers result in an increase of queuing delay and \textit{over-buffering}
\end{itemize}
}
~\\
\only<3->{
\begin{enumerate}
\item[$\Rightarrow$] Buffers should be designed according to their application environment
\end{enumerate}
}
\end{frame}

\begin{frame}{Bufferbloat}
\begin{itemize}
\item Refers to frequently full, excessively large buffers
~\\~\\
\item<2-> Queued packets will not be dropped
	\begin{itemize}
	\item<2->[$\Rightarrow$] Increased congestion, as TCP is not being notified of congestion
	\item<2->[$\Rightarrow$] Decreased network throughput
	\item<2->[$\Rightarrow$] Unnecessary network delay (\textit{latency}), packet delay fluctuation (\textit{jitter})
	\end{itemize}
~\\
\item<3-> With memory getting cheaper over the years, buffers saw an increase in size
\end{itemize}
\end{frame}

@@

		\draw[-, dotted] (net1) -- (Router1);
		\draw[-] (Router1) -- (Router2);
		\draw[-] (Router2) -- (Router3);
		\draw[-, dotted] (Router3) -- (net2);
	\end{tikzpicture}
\end{figure}
~\\
\begin{minipage}{\textwidth}
\begin{columns}
%Content
\only<2->{
	\begin{column}{0.5\textwidth}
		\begin{block}{Transmission Delay}
			Depends on bandwidth and length of the connection
		\end{block}
	\end{column}
}
\only<3->{
	\begin{column}{0.5\textwidth}
		\begin{block}{Processing Delay}
			Time it takes to process a single packet
		\end{block}
	\end{column}
}
%Placeholders
\only<1>{\begin{column}{0.5\textwidth}\end{column}}
\only<1-2>{\begin{column}{0.5\textwidth}\end{column}}
\end{columns}
\end{minipage}

\begin{minipage}{\textwidth}
\begin{columns}
\only<4->{
	\begin{column}{0.5\textwidth}
		\begin{block}{Inter-Packet Delay}
			Allows the receiver to prepare for transmission
		\end{block}
	\end{column}
}
\only<5->{
	\begin{column}{0.5\textwidth}
		\begin{block}{Queuing Delay}
			Time a packet waits in the buffer waiting for transmission
		\end{block}
	\end{column}
}
\only<1-3>{\begin{column}{0.5\textwidth}\end{column}}
\only<1-4>{\begin{column}{0.5\textwidth}\end{column}}
\end{columns}
\end{minipage}
\end{frame}

@@

\begin{column}{0.5\textwidth}
	\begin{block}{Delay}
		\begin{itemize}
		\only<2->{\pgfsetfillopacity{0.2}}
		\item Transmission Delay
		\item Processing Delay
		\item Inter-Packet Delay
		\only<2->{\pgfsetfillopacity{1}}
		\item Queuing Delay
		\end{itemize}
	\end{block}
\end{column}
\end{columns}
~\\~\\
\only<3->{
\begin{itemize}
\item Congestion control algorithms improve
	\begin{itemize}
	\item Queuing delay
	\item Buffer lengths
	\item Buffer management
	\end{itemize}
\end{itemize}
}
\end{frame}

@@

\begin{frame}{TCP}
~\\
\begin{block}{Transmission Control Protocol}
\begin{itemize}
\item Transport protocol
\item Uses IP as internet protocol ($\rightarrow$ TCP/IP)
\item<2-> Ensures successful transmission of all data
\end{itemize}
\end{block}
\begin{itemize}
\item<3-> Transmitted data needs to be acknowledged
\item<3-> Adjusts to the current state of congestion
~\\~\\
\item<4-> Many different implementations exist
\end{itemize}
\end{frame}

@@

\begin{frame}{Perceiving Congestion}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{block}{Round-trip Time (\textit{RTT})}
Time between sending a data packet and receiving its acknowledgment packet
\end{block}
\end{column}
\begin{column}{0.5\textwidth}
\begin{block}{Retransmission Timeout (\textit{RTO})}
Fixed upper limit to RTT
\end{block}
\end{column}
\end{columns}
~\\~\\
\only<2->{
\begin{itemize}
\item Classic TCP perceives congestion by detecting packet loss
\item<3-> Other implementations also perceive congestion by determining variances in RTT over time
\end{itemize}
}
\end{frame}

@@

	\only<1->{\path (s1) edge [above] node {\tiny A} (r1);}
	\only<2->{\path (s2) edge [dashed, above] node {\tiny B} (loss);}
	\only<3->{\path (s3) edge [below] node {\tiny C} (r3);}
	\only<4->{\path (r4) edge [above] node {\tiny A} (s4);}
	\only<5->{\path (r6) edge [above] node {\tiny A,C} (s6);}

@@

\subsection*{Adjusting for Congestion}
\label{sec:ac}
\begin{frame}{Adjusting for Congestion}
~\\
\begin{block}{Goal}
\begin{itemize}
\item Do not fill up buffers in the network
\item Achieve and hold maximum throughput
\item Quickly adapt to congestion
\end{itemize}
\end{block}
\only<2->{
\begin{itemize}
\item Based on Additive-Increase Multiplicative-Decrease principle
\item<3-> $cwnd$ (congestion window) variable
	\begin{itemize}
	\item Current data rate
	\item Maintained by TCP Sender
	\end{itemize}
\end{itemize}
}
\end{frame}

\begin{frame}{Adaption - Additive Increase}
~\\
\begin{block}{Formula}
$cwnd \coloneqq cwnd + u$ for each received ACK packet
\end{block}
\begin{itemize}
\item Linear increase in data rate
\item Slow approach to the networks maximum throughput
\item Prevents the rapid filling of buffers
~\\~\\
\item<2-> Proposed value for $u$: $\frac{1}{cwnd}$
\end{itemize}
\end{frame}

\begin{frame}{Adaption - Slow-Start}
~\\
\begin{block}{Formula}
$cwnd \coloneqq cwnd + 1$ for each received ACK packet
\\$\Rightarrow$ Exponential increase over time
\end{block}
\begin{itemize}
\item Drastic Increase in data rate
\item Quick increase in protocol performance
~\\~\\
\item<2-> Used after
	\begin{itemize}
	\item Connection establishment
	\item Multiple packet losses
	\end{itemize}
\item<3-> Used until
	\begin{itemize}
	\item Slow-Start threshold has been exceeded
	\item Perceived congestion
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Adaption - Multiplicative Decrease}
~\\
\begin{block}{Formula}
$cwnd \coloneqq d \cdot cwnd, 0<d<1$
\end{block}
\begin{itemize}
\item Decreases the data rate drastically
\item Allows congested nodes and buffers to recover
~\\~\\
\item<2-> Proposed value for $d$: $\frac{1}{2}$
\end{itemize}
\end{frame}

@@

\begin{frame}{Network-assisted Congestion Control}
\begin{itemize}
\item Congestion Detection algorithms are very implicit
	\begin{itemize}
	\item Packet loss may also have other causes
	\item No information about the degree of congestion
	\end{itemize}
\item[$\Rightarrow$] Creates unnecessary network delay
~\\~\\
\item<2-> \textit{Explicit congestion notification} (ECN)
	\begin{itemize}
	\item Extension to TCP/IP
	\item Adds \textit{Congestion Experienced} Flag (CE) to the IP header
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Principle of Network-assisted Congestion Control}
\begin{center}
\begin{figure}
	\centering
		\begin{tikzpicture}[->, node distance = 2.1cm]
			%network
			\node[client, scale=0.7] (H1) {};
			\node[router, scale=0.7] (R1) [right=of H1] {};
			\node[router, scale=0.7] (R2) [right=of R1] {};
			\node[router, scale=0.7] (R3) [right=of R2] {};
			\node[client, scale=0.7] (H2) [right=of R3] {};
			\path [line] (H1) -- (R1);
			\path [line] (R1) -- (R2);
			\path [line] (R2) -- (R3);
			\path [line] (R3) -- (H2);
			\only<5->{\path [line, dotted] (H2) -- ++(0,-35pt) -| (H1);}
			\only<5->{\node[below=0.9cm of R2] (text) [transition] {Congested Signal};}

			%packets
			\node[messageclosed,scale=0.5, fill=black!10] (P1) [above right=0.2cm and 1cm of H1] {};
			\only<3->{\node[messageclosed,scale=0.5, fill=black!50] (P2) [above left=0.7cm and 0.3cm of R2] {};}
			\only<2->{\node[messageclosed,scale=0.5, fill=black!10] (P3) [above left=0.4cm and 0.2cm of R2] {};}
			\only<2->{\node[messageclosed,scale=0.5, fill=black!10] (P4) [above left=0.1cm and 0.1cm of R2] {};}
			\only<4->{\node[messageclosed,scale=0.5, fill=black!50] (P5) [above right=0.2cm and 1cm of R3] {};}
			\node[above left = 1.5cm and 0cm of R1] (P) [transition] {Packet};
			\only<2->{\node[above right = 1.5cm and -0.3cm of R2] (CN) [transition] {Congested Node};}
			\only<4->{\node[above left = 1.5cm and -0.4cm of H2] (MP) [transition] {Marked Packet};}
			\path [line] (P) -- (P1);
			\only<2->{\path [line] (CN) -- (R2);}
			\only<4->{\path [line] (MP) -- (P5);}
		\end{tikzpicture}
\end{figure}
\end{center}
\end{frame}

\subsection{XCP}
\label{sec:xcp}
\begin{frame}{XCP}
~\\
\begin{block}{XCP}
\textit{eXplicit Congestion Protocol} generalizes the ECN proposal
\end{block}
\begin{itemize}
\item Introduces new packet control concepts
\item Gives feedback about the degree of congestion
\end{itemize}
~\\
\only<2->{
\begin{block}{XCP Header}
\begin{itemize}
\item $H\_cwnd$
\item $H\_rtt$
\item $H\_feedback$
\end{itemize}
\end{block}
}
\end{frame}

\begin{frame}{XCP Sender}
\begin{itemize}
\item Maintains $cwnd$ and $RTT$
\item Handles packet loss
~\\~\\
\item<2-> On data packet departure
	\begin{itemize}
	\item $H\_cwnd \coloneqq cwnd$
	\item $H\_rtt \coloneqq RTT$
	\item $H\_feedback \coloneqq \frac{r \cdot RTT - cwnd}{cwnd}$, where $r$ is the desired data rate
	\end{itemize}
\item<2-> [$\Rightarrow$] The desired data rate can be reached in just one RTT
~\\~\\
\item<3-> On ACK packet arrival
	\begin{itemize}
	\item $cwnd \coloneqq max(cwnd + H\_feedback, s)$, where $s$ is the packet size
	\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{XCP Router}
\begin{columns}
\begin{column}{.5\textwidth}
\begin{block}{Efficiency Controller (EC)}
\begin{itemize}
\item Maximizes link utilization
\item Minimizes queue lengths and packet drops
\end{itemize}
\end{block}
\end{column}

\begin{column}{.5\textwidth}
\only<2->{
\begin{block}{Fairness Controller (FC)}
\begin{itemize}
\item Divides feedback to all packets fairly
\end{itemize}
\end{block}
}
\end{column}
\end{columns}
~\\
\only<3->{
\begin{itemize}
\item Computed feedback reflects the degree of congestion until this router
\end{itemize}
}
~\\
\only<4->{
\begin{alertblock}{}
Allocates bandwidth equally to all XCP senders
\end{alertblock}
}
\end{frame}

\subsection{Evaluation of Congestion Control}
\label{sec:eval}
\begin{frame}{Evaluation of Congestion Control}
Pros:
\begin{itemize}
\item XCP and ECN outperforms TCP/IP
\end{itemize}
\only<2->{
~\\Cons:
\begin{itemize}
\item Many routers run on hardware only 
	\begin{itemize}
	\item[$\Rightarrow$] Unrecognized packets will be dropped
	\end{itemize}
\item Relies on active queue management
\end{itemize}
}
~\\
\begin{itemize}
\item<3->[$\Rightarrow$] Not used in WANs
\item<4->[$\Rightarrow$] Implementation in SANs more easily
\end{itemize}
\end{frame}

\subsection{DCTP}
\label{sec:dctp}
\begin{frame}{DCTCP}
~\\
\begin{block}{Data center TCP}
\begin{itemize}
\item Extension to TCP for use in data centers
\item Based on ECN
\end{itemize}
\end{block}
\begin{itemize}
\item<2-> Tackles the TCP incast problem
\item<3-> Prevents bursty retransmissions by lowering RTO
\item<3-> Prevents link under-utilization
\begin{itemize}
\item <4-> Adjusts to congestion according to the fraction of packets that experienced congestion
\end{itemize}
~\\~\\
\item<5->[$\Rightarrow$] Implemented in many data centers
\end{itemize}
\end{frame}

@@

\begin{frame}{Conclusion}
\only<1->{
\begin{block}{Origins of Congestion}
\begin{itemize}
\item Buffer size and Buffer management cause delay and packet loss
\item Queuing delay results in increased network delay
\end{itemize}
\end{block}
}
\only<2->{
\begin{block}{Congestion Avoidance}
\begin{itemize}
\item Simple way to tackle congestion
\item Only gets implicit congestion notification
\item Often causes instead of resolving network stress
\end{itemize}
\end{block}
}
\only<3->{
\begin{block}{Network-assisted Congestion Control}
\begin{itemize}
\item Can explicitly control congestion
\item Results in overall better network performance
\item Implementations in WANs very challenging, successful in data centers
\end{itemize}
\end{block}
}
\end{frame}
